---
title: "Bayes' Rule - Magician Michelles's Prediction of Megathrust Earthquakes"
author: "Gwendolin Lüdtke, Sina Spors"
date: "27 November 2018"
output: html_document
#html_document:
 #   highlight: textmate
  #  code_folding: show
   # theme: spacelab                 
    #toc: true 
    #toc_float: 
     # collapsed: false 
      #smooth_scroll: true 
runtime: shiny

#<style>
#body {
#text-align: justify} 
#</style>

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Magician Michelle says that she can predict large megathrust earthquakes. Well, at least 95 % of the time. Assume that these earthquakes occur with a probability of 10 <sup>-5</sup>. How likely will Michelle cause a false alarm with her next prediction?*

## Recap: Classic RMarkdown HTML
In order to calculate the probabiliy of a false alarm regarding large megathrust earthquakes, the **Bayes' Rule** should be applied:

$$p(\kappa|D)=\frac{p(D|\kappa) p(\kappa)}{p(D)}$$
where

* D = Michelle claims that she can predict megathrust earthquakes 
* $\kappa$ = occurence of megathrust earthquakes 

We can deduce from the values given in the task that:

* $p(D|\kappa)=p(Michelle\:is\:right)=95/100$
* $p(\kappa)=p(megathrust\:earthquake)=1/100000$

In R:

```{r}
#p(k) = p(megathrust earthquake)
prior <- 1/100000

#p(D|k) = p(Michelle says that there is a megathrust earthquake|megathrust earthquake)
likeli <- 95/100
```

Again, we are using the **law of total probability** in order to calculate $p(D)$:

$$p(D) = p(D|\kappa)p(\kappa)+p(D|-\kappa)p(-\kappa)$$

```{r}
#p(D) = p(Michelle says that there is a megathrust earthquake)
evid <- (likeli*prior) + ((1-likeli)*(1-prior))
```

```{r, echo = FALSE}
#echo = FALSe, do not show chunk, but output
cat("p(D): ",evid)

```

## Final Calculation

```{r}
#p(k|D) = p(megathrust earthquake|Michelle says that there is a megathrust earthquake)
post <- likeli * prior / evid
```

```{r, echo = FALSE}
cat("p(k|D): ",post)
```

This is the probability of Michelle being right about the occurence of a megathrust earthquake.

## Shiny 

However, we are looking for the probability of a false alarm. That is why we have to calculate the complementary probability: 

```{r}
post_false_alarm <- 1 - post
```

```{r, echo = FALSE}
cat("p(k|D): ",post_false_alarm)
```

Don't trust Michelle! 

This was a classic RMarkdown Document.

Now, we can change the input parameters (probability of Michelle being right and probability of a megathrust earthquake) in order to see the changes in the posterior. 

The user should use a slider input / ```sliderInput``` in order to change the ranges of the parameters. This kind of input is advantageous since the programmer can define/set the minium and maximum value as well as the steps. 


```{r, echo=FALSE}
sliderInput("trust", "Trustworthiness (in %):",95, min = 1, max = 100)
sliderInput("earthquake", "Probability of a megathrust earthquake (in %):",1, min = 0.0000001, max = 100)

renderText({
  likeli <- input$trust
  prior <- input$earthquake
  evid <- (likeli*prior) + ((100-likeli)*(100-prior))
  post <- (likeli * prior / evid) * 100
  
  #format(post, digits = 2 )
  sprintf("The trustworthiness is: %.2f %%", post)
  #paste("The trustworthiness is:", post, " %")
})

renderPlot({
  likeli <- input$trust
  prior <- input$earthquake
  evid <- (likeli*prior) + ((100-likeli)*(100-prior))
  post <- (likeli * prior / evid) * 100
  
  
  barplot(post,main = "",horiz = TRUE, xlim = c(0, 100), col = c("#4286f4"), names.arg = c("trustworthiness"))
})





#renderPlot({

 # plot(input$trust, input$earthquake)
#})

#renderText({
#  likeli <- input$trust
#  prior <- input$earthquake
#  evid <- (likeli*prior) + ((1-likeli)*(1-prior))
#  post <- likeli * prior / evid
#  
#  paste("The trustworthiness is:", post)
#  })

```



