---
title: "Bayes' Rule - Magician Michelles's Prediction of Megathrust Earthquakes"
author: "Gwendolin Lüdtke, Sina Spors"
date: "27 November 2018"
output: html_document
#html_document:
 #   highlight: textmate
  #  code_folding: show
   # theme: spacelab                 
    #toc: true 
    #toc_float: 
     # collapsed: false 
      #smooth_scroll: true 
runtime: shiny

#<style>
#body {
#text-align: justify }
#</style>



---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<div style="text-align: justify;">
*Magician Michelle says that she can predict large megathrust earthquakes. well, at least 95 % of the time. Assume that these earthquakes occur with a probability of 10<sup> -5</sup>. How likely will Michelle cause a false alarm with her next prediction?*

Up to this point, a classic R Markdown HTML document was used for documenting possible applications of the Bayes' Rule. Data which was used in the calculation was already provided by the programmer and the reader could only track the different steps of calculation. Now, we go one step further and include a user input in our calculation of a probability by using the Bayes' Rule. For this reason, we are using a Shiny document.
</div>

## Recap: Classic R Markdown HTML

<div style="text-align: justify;">
As for the wildfire-problem, the **Bayes' Rule** is applied in order to calculate the probability of a false alarm regarding large mega-thrust earthquakes.

$$p(\kappa|D)=\frac{p(D|\kappa) p(\kappa)}{p(D)}$$
where

* D = Michelle predicts a mega-thrust earthquakes 
* $\kappa$ = a mega-thrust earthquake occurs

We can deduce from the values given in the task that:

* $p(D|\kappa)=p(Michelle\:is\:right)=95/100$
* $p(\kappa)=p(mega thrust\:earthquake)=1/100000$

In R:
</div>

```{r}
#p(k) = p(megathrust earthquake occurs)
prior <- 1/100000

#p(D|k) = p(Michelle predicts a megathrust earthquake|megathrust earthquake occurs)
likeli <- 95/100
```

<div style="text-align: justify;">
Again, we are using the **law of total probability** in order to calculate $p(D)$:
</div>

$$p(D) = p(D|\kappa)p(\kappa)+p(D|-\kappa)p(-\kappa)$$

```{r}
#p(D) = p(Michelle says that there is a megathrust earthquake)
evid <- (likeli*prior) + ((1-likeli)*(1-prior))
```

```{r, echo = FALSE}
#echo = FALSE, do not show chunk, but output
cat("p(D): ",evid)

```

<div style="text-align: justify;">
Finally, the probability of Michelle being right about the occurrence of a mega-thrust earthquake can be calculated:
</div>

```{r}
#p(k|D) = p(megathrust earthquake|Michelle says that there is a megathrust earthquake)
post <- likeli * prior / evid
```

```{r, echo = FALSE}
cat("p(k|D): ",post)
```

<div style="text-align: justify;">
However, we are looking for the probability of a false alarm. That is why we have to calculate the complementary probability: 
</div>

```{r}
post_false_alarm <- 1 - post
```

```{r, echo = FALSE}
cat("p(k|D): ", post_false_alarm)
```

Don't trust Michelle! 

## Shiny R Markdown Documents

<div style="text-align: justify;">
https://bookdown.org/yihui/rmarkdown/shiny-documents.html

In contrast to the prior R Markdown HTML, Shiny documents are no longer static documentations. They allow that the user can change parameters of the calculation/analysis and provide a immediate output of the results. 

Instead of "embed a standalone Shiny application within a document", we decided to embed "individual Shiny inputs and outputs in R Markdown".
</div>

### Defining the Layout

<div style="text-align: justify;">
First of all, we define the layout of the Shiny App. There are different layout options. We are using the function ```sidebarLayout()``` to create an app-layout with a sidebar that contains the input controls and a larger main area that contains the outputs.
</div>

```{r, eval = FALSE}
sidebarLayout(
  sidePanel(
    
  ),
  mainPanel(
    
  )
)

```

### Selection of Parameters

<div style="text-align: justify;">
With the help of Shiny functions (?), the user can now change/select the input parameters:
* probability of Michelle being right
* probability of a mega-thrust earthquake

... and can immediately see the changes in the posterior-probability. 

The user should use a ```sliderInput``` in order to change the values of the input parameters. This kind of input is advantageous since the programmer can define the minimum and maximum values as well as the steps.
</div>

###Output

<div style="text-align: justify;">
It is important to use ```renderPlot``` to create [flexible graphics). 
</div>

```{r}

# Title of the app
titlePanel("Bayes' Rule")

sidebarLayout(
  
  sidebarPanel( # For the input controls
    
    sliderInput("trust", "Trustworthiness of Michelle:",min = 1, max = 100, value = 95, step = 1, 
                  post= " %",
                  animate = TRUE
    ),
    
    sliderInput("earthquake", "Probability of a megathrust earthquake:", 
                min = 0.0000001, max = 100, value = 0.00001, step = 1,
                post = " %",
                animate = TRUE)

    ),
  
  mainPanel( # for the outputs

   renderText({
   
      likeli <- (input$trust)/100 #divide by 100 to work with xxx
      prior <- (input$earthquake)/100
      evid <- (likeli*prior) + ((1-likeli)*(1-prior))
      post <- likeli * prior / evid
  
      #paste("Michelle is right with a probability of ", post, " %")
     #paste("Michelle is wrong and causes false alarm with a probability of ", (1-post), " %")
     postPercentage <- (1-post) * 100
     sprintf("Michelle is wrong and causes false alarm with a probability of %.4f %%", postPercentage)
      #cat (paste (post, "\n", collapse=""))
      #format(post, digits = 2 )
     # sprintf("Michelle is right with a probability of: %.2f %%", post)
    }),
   
   renderPlot({

      likeli <- (input$trust)/100
      prior <- (input$earthquake)/100
      evid <- (likeli*prior) + ((1-likeli)*(1-prior))
      post <- likeli * prior / evid
      
      postPercentage <- (1-post) * 100
      barplot(postPercentage,main = "",horiz = TRUE, xlim = c(0, 100), col = c("#4286f4"), names.arg =  c("trustworthiness"))

      #barplot((1-post),main = "",horiz = TRUE, xlim = c(0, 1), col = c("#4286f4"), names.arg =  c("trustworthiness"))
  })
    
  )
)

```

[das kann nachher alles raus]
## try

see comments on sidebarLayout

use of sliderInput / creation of slider controls

```{r, echo=FALSE, eval=FALSE}

      # Input: Specification of range within an interval ----
      sliderInput("range", "Range:",
                  min = 1, max = 1000,
                  value = c(200,500)),

      # Input: Custom currency format for with basic animation ----
      sliderInput("format", "Custom Format:",
                  min = 0, max = 10000,
                  value = 0, step = 2500,
                  pre = "$", sep = ",",
                  animate = TRUE),

    
  ),
  
  mainPanel(
    # Output: Table summarizing the values entered ----
     # tableOutput("values")
  )
  
  
)


```


use the function selectInput to creat a box with choises to select from (widgets)

after the selection of single value or range > slider Input anpassen (use of conditionalPanel)

```{r, eval=FALSE}

selectInput("select","Select: Single Value or Range?",
            c("Single Value" ="sin",
              "Range" ="rang"),
            multiple = FALSE
            )


#only show this panel if single values are selected
conditionalPanel(
  condition = "input.select == 'sin'",
  sliderInput(
    "trust",
    "Trustworthiness (in %):",95, min = 1, max = 100
  ),
  sliderInput("earthquake", "Probability of a megathrust earthquake (in %):",0.00001, min = 0.0000001, max = 100),
  plotOutput("scatterPlot", height = 300)
  
)

conditionalPanel(
  condition = "input.select == 'rang'",
  sliderInput(
    "trust",
    "Trustworthiness (in %):", min = 1, max = 100, value=c(5,10)
  ),
  sliderInput("earthquake", "Probability of a megathrust earthquake (in %):",min = 0.0000001, max = 100,
  value =c(1,10))
)

output$scatterPlot <- renderPlot({
    x <- input$trust
    y <- input$earthquake
    plot(x, y)
})

```

